Concentration of Maxima and Fundamental Limits in High-Dimensional Testing and Inference

### 2021-01-15 4.exact_support_recovery/4.minimax.tex
updated Ch 4
I updated up to and including Ch 4.2.2.

Hope to finish tomorrow.

Meanwhile, if we search over \emph{all procedures}, but expand the parameter space to include {all} dependence structures, then a different minimax optimality statement holds for Bonferroni's procedure.
This result, formally stated in Section \ref{subsec:minimax-over-dependence}, is a consequence of our characterization of the finite-sample Bayes optimality of thresholding procedures in Sections \ref{subsec:Bayes-optimality} and \ref{subsec:optimal-procedure-sub-exponential}.
% The result is of independent interest and it is the main contribution of this section.

@@

In studying support recovery problems (e.g., \citet{arias2017distribution}), restrictions to the thresholding procedures are sometimes justified by arguing that such procedures are the ``reasonable'' choice for estimating the support set.
%  In Theorem \ref{thm:sufficient} we saw that a (FWER-controlling) thresholding procedure is indeed consistent for support recovery. 
% It is natural to ask if it will be sufficient to restrict our attention to only such procedures.
We show in this section that, perhaps surprisingly, for general error models, thresholding procedures are not always optimal, even when the observations are independent.

We shall identify the optimal procedure for support recovery problems under a Bayesian setting with general distributional assumptions (including but not limited to additive models \eqref{eq:model-additive}).
Specifically, we assume that there is an ordered set $P = (i_1,\ldots,i_s)$, $i_i\in\{1,\ldots,p\}$, and $s$ (not necessarily equal) densities $f_{1}, \ldots, f_{s}$, such that the observations indexed by $P$ have corresponding densities. That is,
\begin{equation} \label{eq:signal-distributions-ordered}
x(i_j) \sim f_j, \quad j=1,\ldots,s.
\end{equation}
Let also the rest $(p-s)$ observations have common density $f_0$, i.e., $x(i)\sim f_0$ for $i\not\in S$.
We further assume that the observations $x$ are mutually independent.

We adopt here a Bayesian framework to measure statistical risks. 
Let the ordered support $P=(i_1,\ldots,i_s)$ have prior
\begin{equation} \label{eq:uniform-ordered}
\pi((i_1,\ldots, i_s)) = {(p-s)!}/{p!},
\end{equation}
for all distinct $1\le i_1, \ldots, i_s\le p$.
Consequently, the unordered support $S=\{i_1,\ldots,i_s\}$ is distributed uniformly in the collection of all set of size $s$, with the unordered uniform distribution $\pi^{\text{u}}$. That is, for all for all $S\in\mathcal{S}:=\left\{S\subseteq\{1,\ldots,p\};|S|=s\right\}$, we have 
\begin{equation} \label{eq:uniform}
\pi^{\text{u}}
(\{i_1,\ldots, i_s\}) = \sum_{\sigma}\pi((i_{\sigma(1)},\ldots, i_{\sigma(s)})) = {(p-s)!s!}/{p!},
\end{equation}
where the sum is taken over all permuations of $\{1,2,\ldots,s\}$.

For any fixed configuration $P$, consider the loss function,
$$
\ell(\widehat{S},S) := \P[\widehat{S} \neq S] = \P_{P}[\widehat{S} \neq S],
$$
where the probability is taken over the randomness in the observations only.
The Bayes optimal procedures should minimize 
\begin{equation} \label{eq:Bayes-risk}
    \E_\pi \P[\widehat{S} \neq S],
\end{equation}
where the expectation is taken over the random configurations $P$, with a uniform distribution $\pi$ as specified in \eqref{eq:uniform-ordered}.

@@

Their relationship is summarized in the following lemma.

>>>>

Meanwhile, if we search over \emph{all procedures}, but expand the space of models to include {\em all} dependence structures, then a different minimax optimality statement holds for Bonferroni's procedure.
This result, formally stated in Section \ref{subsec:minimax-over-dependence}, is a consequence of our characterization of the finite-sample Bayes optimality of thresholding procedures in Sections \ref{subsec:Bayes-optimality} and \ref{subsec:optimal-procedure-sub-exponential}.
% The result is of independent interest and it is the main contribution of this section.

@@

\stilian{\fbox{SUGGESTION}}{ What do you think about making the Bayesian stuff a Chapter?  I think it stands well on its own... plus it is somewhat different from the ``non-Bayesian'' setup of the previous support recovery results...}

\stilian{
In studying support recovery problems \citep[e.g.,][]{arias2017distribution}, often the attention is restricted to the class of
thresholding procedures.  This is justified by arguing that such procedures are the ``reasonable'' choice for estimating the support set.
%  In Theorem \ref{thm:sufficient} we saw that a (FWER-controlling) thresholding procedure is indeed consistent for support recovery. 
% It is natural to ask if it will be sufficient to restrict our attention to only such procedures.
We show in this section that, in many cases, the finite-sample Bayes optimal procedures do indeed belong to the class of thresholding procedures.  
Perhaps surprisingly, however, we also show that for general error models, the thresholding procedures are sub-optimal, 
even when the observations are independent.  Roughly speaking, in the additive error 
models \eqref{eq:model-additive},  this distinction depends on whether or not the error density is log-concave 
(Proposition \ref{prop:log-concave} and Corollary \ref{cor:log-convex}).}{\fbox{edited, agree?}}

We shall identify the optimal procedure for support recovery problems under a Bayesian setting with general distributional assumptions 
(including but not limited to additive models \eqref{eq:model-additive}). 
Specifically, consider an ordered set $P = (i_1,\ldots,i_s)$ of $s$ distinct indices in $\{1,\ldots,p\}$, which will be referred to 
as the {\em ordered support} of the signal.  Consider also $s$ (not necessarily equal) signal densities 
$f_{1}, \ldots, f_{s}$ as well as an error density $f_0$. We shall assume that the observations with indices $i_j$ 
are drawn independently from $f_j$:
\begin{equation} \label{eq:signal-distributions-ordered}
x(i_j) \sim f_j, \quad j=1,\ldots,s,
\end{equation}
while the remaining $(p-s)$ observations have common density $f_0$, i.e., $x(i)\sim f_0$ for $i\not\in S$. 
We further assume that the observations $x$ are mutually independent.

We adopt now a Bayesian framework to measure statistical risks and assume that the underlying (unknown) support is {\em random}. Namely,
let the ordered support $P=(i_1,\ldots,i_s)$ have prior
\begin{equation} \label{eq:uniform-ordered}
\pi((i_1,\ldots, i_s)) = {(p-s)!}/{p!},
\end{equation}
for all distinct $1\le i_1, \ldots, i_s\le p$.
Consequently, the unordered support $S=\{i_1,\ldots,i_s\}$ is distributed uniformly over the collection of all sets of size $s$, with 
the unordered uniform distribution $\pi^{\text{u}}$. That is, for all $S\in\mathcal{S}:=\left\{S\subseteq\{1,\ldots,p\};|S|=s\right\}$, we have 
\begin{equation} \label{eq:uniform}
\pi^{\text{u}}
(\{i_1,\ldots, i_s\}) = \sum_{\sigma}\pi((i_{\sigma(1)},\ldots, i_{\sigma(s)})) = {(p-s)!s!}/{p!},
\end{equation}
where the sum is taken over all permuations of $\{1,2,\ldots,s\}$.

For any fixed configuration $P$, consider the loss function,
$$
\ell(\widehat{S},S) := \P[\widehat{S} \neq S] = \P_{P}[\widehat{S} \neq S],
$$
where the probability is taken over the randomness in the observations only. 
\stilian{(The $x(i)$'s are conditionally independent, given the ordered support $P$.)}{\fbox{added.  Agree?}}
The Bayes optimal procedures should minimize 
\begin{equation} \label{eq:Bayes-risk}
    \E_\pi \P[\widehat{S} \neq S],
\end{equation}
where the expectation is taken over the random configurations $P$, with a uniform distribution $\pi$ as specified in \eqref{eq:uniform-ordered}.

@@

The relationship between the above two notions is summarized next.

### 2021-01-15 4.exact_support_recovery/4.numerical.tex
updated Ch 4
I updated up to and including Ch 4.2.2.

Hope to finish tomorrow.

We examine numerically the boundaries \eqref{eq:strong-classification-boundary} under several error tail assumptions for independence errors in this section.
Numerical experiments for dependent errors will be deferred until we characterize the \ac{URS} conditions in Chapter \ref{chap:URS}.

To demonstrate the phase transition phenomenon under different error tail densities, we simulate from the additive error model \eqref{eq:model-additive} with
\begin{itemize}
    \item Gaussian errors, where the density is given by
    $f(x) = \frac{1}{\sqrt{2\pi}}\exp{\left\{-x^2/2\right\}}$.
    \item Laplace errors, where the density is given by $f(x) = \frac{1}{2}\exp{\left\{-\left|x\right|\right\}}$.
    \item Generalized Gaussian $\nu=1/2$, with density
    $f(x) = \frac{1}{2}\exp{\big\{-2\left|x\right|^{1/2}\big\}}$.
\end{itemize}
The sparsity and signal size of the sparse mean vector are parametrized as in equations \eqref{eq:sparsity-parametrized} and \eqref{eq:signal-size-parametrized}, respectively.
The support set $S$ is estimated with 
$\widetilde{S} = \left\{i:x(i)>\sqrt{2\log{p}}\right\}$ 
under the Gaussian errors, 
$\widetilde{S} = \left\{i:x(i)>\log{p} + (\log{\log{p}})/2\right\}$ 
under the Laplace errors, and with
$\widetilde{S} = \{i:x(i)> \frac{1}{4}\left(W\left(-c/(ep\log{p})\right) + 1\right)^2\}$
under the generalized Gaussian ($\nu = 1/2$) errors. Here $W$ is the Lambert W function, i.e., $W=f^{-1}$ where $f(x)=x\exp{(x)}$.
The choices of thresholds correspond to Bonferroni's procedures with FWER decreasing at a rate of $1/\sqrt{\log{p}}$, therefore satisfying the assumptions in Theorem \ref{thm:sufficient}.
Experiments were repeated 1000 times under each sparsity-and-signal-size combination.

The results of the numerical experiments are shown in Figure \ref{fig:phase-simulated}.
The numerical results illustrate that the predicted boundaries are not only accurate in high-dimensions ($p=10000$, right panels of Figure \ref{fig:phase-simulated}), but also practically meaningful even at moderate dimensions ($p=100$, left panels of Figure \ref{fig:phase-simulated}).

\begin{figure}
      \centering
      \includegraphics[width=0.4\textwidth]{./figures/simulated_phase_diagram_p100.eps}
      \includegraphics[width=0.4\textwidth]{./figures/simulated_phase_diagram_p10000.eps}
      \includegraphics[width=0.4\textwidth]{./figures/simulated_phase_diagram_Laplace_p100_4.eps}
      \includegraphics[width=0.4\textwidth]{./figures/simulated_phase_diagram_Laplace_p10000_4.eps}
      \includegraphics[width=0.4\textwidth]{./figures/simulated_phase_diagram_NLC_p100.eps}
      \includegraphics[width=0.4\textwidth]{./figures/simulated_phase_diagram_NLC_p10000.eps}
      \caption{The empirical probability of exact support recovery from numerical experiments, as a function of sparsity level $\beta$ and signal sizes $r$, from Gaussian error models (upper panels), Laplace error models (middle panels), and generalized Gaussian with $\nu=1/2$ (lower panels); darker color indicates higher probability of exact support recovery. 
      The experiments were repeated 1000 times for each sparsity-signal size combination, and for dimensions $p=100$ (left panels) and $p=10000$ (right panels). Numerical results agree with the boundaries described in Theorem \ref{thm:sufficient}; convergence is noticeably slower for under generalized Gaussian ($\nu=1/2$) errors.
      For reference, the dashed and dash-dotted lines represent the weak classification and detection boundaries (see Chapter \ref{chap:phase-transitions}).}
      \label{fig:phase-simulated}
\end{figure}

>>>>

\stilian{Here, we examine numerically the applicability of the asymptotic boundaries 
\eqref{eq:strong-classification-boundary} for finite $p$.  We consider independent errors for several popular models of the marginal distribution.  Numerical experiments for dependent errors will be deferred to Chapter \ref{chap:URS}.}{\fbox{check}}

To demonstrate the phase transition phenomenon under different error tail densities, we simulate from the additive error model \eqref{eq:model-additive} with
\begin{itemize}
    \item Gaussian errors, where the density is given by
    $f(x) = \frac{1}{\sqrt{2\pi}}\exp{\left\{-x^2/2\right\}}$.
    \item Laplace errors, where the density is given by $f(x) = \frac{1}{2}\exp{\left\{-\left|x\right|\right\}}$.
    \item Generalized Gaussian $\nu=1/2$, with density
    $f(x) = \frac{1}{2}\exp{\big\{-2\left|x\right|^{1/2}\big\}}$.
\end{itemize}
The sparsity and signal size of the sparse mean vector are parametrized as in equations \eqref{eq:sparsity-parametrized} and \eqref{eq:signal-size-parametrized}, respectively.
The support set $S$ is estimated with 
$\widetilde{S} = \left\{i:x(i)>\sqrt{2\log{p}}\right\}$ 
under the Gaussian errors, 
$\widetilde{S} = \left\{i:x(i)>\log{p} + (\log{\log{p}})/2\right\}$ 
under the Laplace errors, and with
$\widetilde{S} = \{i:x(i)> \frac{1}{4}\left(W\left(-c/(ep\log{p})\right) + 1\right)^2\}$
under the generalized Gaussian ($\nu = 1/2$) errors. Here $W$ is the Lambert W function, i.e., $W=f^{-1}$ where $f(x)=x\exp{(x)}$.
The choices of thresholds correspond to Bonferroni's procedures with FWER decreasing at a rate of 
${\cal O}(1/\sqrt{\log{p}})$, therefore satisfying the assumptions in Theorem \ref{thm:sufficient}.
The experiments were repeated independently 1000 times under each sparsity-and-signal-size combination.

\stilian{Figure \ref{fig:phase-simulated} shows the resulting empirical probabilities of exact support recovery over 
a grid of $r$ and $\beta$ values for the cases of small dimensions ($p=100$, left panels) and high dimensions ($p=10\ 000$, right panels).  
Observe that the asymptotic boundary is rather accurate in the high-dimensional regime.  
The phase-transition phenomenon is also evident and practically relevant in dimensions as small 
as $p=100$.}{ \fbox{suggested replacement of the following}
 The results of the numerical experiments are shown in Figure \ref{fig:phase-simulated}.
The numerical results illustrate that the predicted boundaries are not only accurate in high-dimensions ($p=10^4$, right panels of Figure \ref{fig:phase-simulated}), but also practically meaningful even at moderate dimensions ($p=100$, left panels of Figure \ref{fig:phase-simulated}).}

     \stilian{\fbox{Question:}} {In Figure \ref{fig:phase-simulated}
      Do we know what the DETECTION boundary is for Non-Gaussian  AGG$(1/2)$?  You are plotting a line in the 
      bottom panels for the DETECTION boundary, but is this exactly a line?  Has this been studied in the literature 
      actually or are we CONJECTURING this boundary -- note that Chapter 3 focuses on the Gaussian case only and we don't prove the detection 
      boundary, right?}

\begin{figure}
      \centering
      \includegraphics[width=0.4\textwidth]{./figures/simulated_phase_diagram_p100.eps}
      \includegraphics[width=0.4\textwidth]{./figures/simulated_phase_diagram_p10000.eps}
      \includegraphics[width=0.4\textwidth]{./figures/simulated_phase_diagram_Laplace_p100_4.eps}
      \includegraphics[width=0.4\textwidth]{./figures/simulated_phase_diagram_Laplace_p10000_4.eps}
      \includegraphics[width=0.4\textwidth]{./figures/simulated_phase_diagram_NLC_p100.eps}
      \includegraphics[width=0.4\textwidth]{./figures/simulated_phase_diagram_NLC_p10000.eps}
      \caption{The empirical probability of exact support recovery from numerical experiments, as a function of sparsity level $\beta$ and signal sizes $r$, 
      from Gaussian error models (upper panels), Laplace error models (middle panels), and generalized Gaussian with $\nu=1/2$ (lower panels); darker color indicates higher probability of exact support recovery. 
      The experiments were repeated 1000 times for each sparsity-signal size combination, and for dimensions $p=100$ (left panels) and $p=10000$ (right panels). 
      The numerical results agree with the boundaries described in Theorem \ref{thm:sufficient}. The convergence is noticeably slower for under the heavier 
      generalized Gaussian ($\nu=1/2$) errors.  For reference, the dashed and dash-dotted lines represent the weak classification and detection boundaries
       (see Chapter \ref{chap:phase-transitions}).
      \label{fig:phase-simulated}}
\end{figure}

### 2021-01-15 5.URS/URS.tex
minor edits
added

1) ?.Example.tex files for each chapter conforming to the current style.

2) Began the "index".  If you want to include a term do \index{term}{term}

3) Fixed the level of the table-of-contents.
updated Ch 4
I updated up to and including Ch 4.2.2.

Hope to finish tomorrow.

Suppose that the array of errors  ${\cal E}$ comes from a stationary Gaussian time series $\epsilon(i),\ i\in \mathbb{N}$, with auto-covariance $r_p=\cov(\epsilon(i+p),\epsilon(i))$. 
One is interested in the asymptotic behavior of the maxima $M_p:=\max_{i=1,\dots,p} \epsilon(i)$.

>>>>

\medskip
\noindent{\bf The Berman and UDD conditions.} 
 \label{UDD_and_Berman}
 Suppose that the array of errors  ${\cal E}$ comes from a stationary Gaussian time series $\epsilon(i),\ i\in \mathbb{N}$, with auto-covariance $r_p=\cov(\epsilon(i+p),\epsilon(i))$. 
One is interested in the asymptotic behavior of the maxima $M_p:=\max_{i=1,\dots,p} \epsilon(i)$.
